### 2017-02-16 chapters/ch2-checkpoint.tex
Write some validation proofs

\subsection{Notation and definition}
\begin{itemize}

@@

\item Given two TX blocks $t_{i,j}$ and $t_{i',j'}$, but $i \neq i'$. If $h_s =
  h'_s$, $h_r = h'_r$, $m = m'$ and the signatures are valid, then we call them
  a \emph{pair}.

@@

\item The result of a consensus in round $r$ is a set of two-tuple of CP.
  Namely, $\mathcal{C}_r = \{ (c_{i,j}, c_{i,k}) : j < k, \text{agreed by the
    promoters}\}$.
\end{itemize}

\subsection{Properties}
\begin{itemize}
\item A TX block has two \emph{ancestor} blocks. Let a pair be $t_{i,j}$ and
  $t_{i',j'}$, then $(b_{i, j-1}, b_{i', j'-1})$ is the input block of $t_{i,j}$.

@@

\item Every TX block $t_{i,j}$ is enclosed by two CP blocks $(c_{i,a},
  c_{i,b})$, where $a = \argmin_{k, k < j, \texttt{typeof}(b_{i,k}) = \gamma}(j
  - k)$ and $b = \argmin_{k, k > j, \texttt{typeof}(b_{i,k}) = \gamma}(k - j)$.

\item The blocks between two $\gamma$ blocks $(c_{i,a}, c_{i,b})$ is $\{b_{i,j}
  : b_{i,j} \in B_i, a \le j \le b\}$.

@@

\subsection{Luck value}
First we define the luck value $l_{i,j} = \texttt{h}(k_i || c_{i,j})$, where $k$ is the
public key of $i$. A lower luck value equates to higher luck.

@@

\begin{lemma}
  If a node sees a valid $\mathcal{C}_{r-1}$ and another node sees a valid
  $\mathcal{C}'_{r-1}$, then $\mathcal{C}_{r-1} = \mathcal{C}'_{r-1}$.
\end{lemma}
\begin{proof}
  (sketch) $\mathcal{C}_{r-1}$ is signed by at least $N-f$ promoters from round
  $r-1$.
\end{proof}

@@

\begin{lemma}
  The new set of promoter for the next consensus round is consistent with
  respect to all the nodes in the network.
\end{lemma}
\begin{proof}
  (sketch) All nodes use the same deterministic function to compute the luck
  value.
\end{proof}

Nodes should wait for some time to collect the CP blocks, so they wait for some
time $\Delta$ before moving on to the next phase.

\begin{lemma}
  Promoters waiting for a some time $\Delta$ to collect transactions does not
  violate the asynchronous assumption.
\end{lemma}
\begin{proof}
  (sketch) This can be seen as a long delay in the asynchronous system.
\end{proof}

\begin{corollary}
  The setup phase satisfies the validity, agreement and liveness properties.
\end{corollary}
\begin{proof}
  (sketch) Validity and agreement is satisfied because the promoters are
  computed using a deterministic algorithm using the latest consensus result.
  Liveness is satisfied because $\mathcal{C}_{r-1}$ is eventually propagated to
  all node by gossiping.
\end{proof}

@@

There techniques that uses no dealers. First is to use
PBFT~\cite{castro1999practical}, but we must change our asynchronous assumption
into the weak synchrony assumption. Most likely this is not possible because of
the ``When to start?'' problem. It's difficult to give a bounded delay for
propagating the consensus result.

@@

\section{Fraud detection}

\subsection{Threat model}
\begin{itemize}
\item What is the influence of a fork for future transactions?
\item What if both parties (sender and receiver) are malicious?
\end{itemize}

\subsection{Random sampling}

>>>>

\subsection{Notation, definition and properties}
\begin{itemize}

@@

\item Given two TX blocks $t_{i,j}$ and $t_{i',j'}$, if $i \ne i'$, $i = s$, $i'
  = r$, $h_s = h'_s$, $h_r = h'_r$, $m = m'$ and the signatures are valid, then
  we call them a \emph{pair}. Note that given one TX block, the pair can be
  determined directly.

@@

\item We define
  $$\texttt{round}: C_i \rightarrow \mathbb{R}_{\ge 1}$$
  as a function that gets the consensus round number used to create the given CP
  block.

\item The CP blocks that follows a pair of TX blocks must be created using the
  same $\mathcal{C}_r$, otherwise the transaction is invalid. Concretely, given
  a pair $t_{i,j}$ and $t_{i',j'}$, then there exist $c_{i, k}$ and $c_{i', k'}$
  where $j < k$, $j' < k'$, $\{ \texttt{typeof}(b_{i, x}) : x \in \{j, \dots k -
  1\}\}$ are all $\tau$, $ \{ \texttt{typeof}(b_{i', x}) : x \in \{ j', \dots k'
  - 1\} \}$ are all $\tau$ and $\texttt{round}(c_{i, k}) = \texttt{round}(c_{i',
    k'})$. This constraint is not the result of the consensus protocol, but the
  validation protocol.

\item The result of a consensus in round $r$ is a set of CP blocks. Namely,
  $$\mathcal{C}_r = \{ c_{i,j} : \text{agreed by the promoters}\}.$$

\item A TX block can be valid, invalid or unknown. All TX blocks begin as
  unknown, they can be validated using our validation protocol.

\item A TX block has two \emph{ancestor} blocks. Let a pair be $t_{i,j}$ and
  $t_{i',j'}$, then $(b_{i, j-1}, b_{i', j'-1})$ is the input block of $t_{i,j}$.

@@

\item Every TX block $t_{i,j}$ is enclosed by two CP blocks $(c_{i,a},
  c_{i,b})$, where
  $$a = \argmin_{k, k < j, \texttt{typeof}(b_{i,k}) = \gamma}(j - k),$$
  $$b = \argmin_{k, k > j, \texttt{typeof}(b_{i,k}) = \gamma}(k - j).$$
  We call this the \emph{enclosure} of $t_{i,j}$.

\item The \emph{piece} of $t_{i,j}$ defined by the enclosure $(c_{i,a},
  c_{i,b})$ is $\{b_{i,j} : a \le j \le b\}$.
  We define
  $$ \texttt{pieces}: B_i \rightarrow P(B_i) $$
  as the function that returns the pieces ($P$ denotes power set).

\item Every TX block $t_{i,j}$ is enclosed by two \emph{agreed} CP blocks
  $(c_{i,a}, c_{i,b})$ (CP blocks that are in some consensus result), where
  $$a = \argmin_{k, k < j, \texttt{typeof}(b_{i,k}) = \gamma}(j - k),$$
  $$b = \argmin_{k, k > j, \texttt{typeof}(b_{i,k}) = \gamma}(k - j).$$
  We call this the \emph{agreed enclosure} of $t_{i,j}$.

\item The \emph{agreed piece} of $t_{i,j}$ defined by the agreed enclosure
  $(c_{i,a}, c_{i,b})$ is $\{b_{i,j} : a \le j \le b\}$. Note that \emph{piece}
  $\subseteq$ \emph{agreed piece} for some TX block.
  We define
  $$ \texttt{a-pieces}: B_i \rightarrow P(B_i) $$
  as the function that returns the agreed pieces.

@@

\subsection{Luck value}
First we define the luck value $l_{i,j} = \texttt{h}(k_i || c_{i,j})$, where $k$
is the public key of $i$. A lower luck value equates to higher luck. We assume
an application agnostic system and do not attempt to defend against the sybil
attack.

An alternative is to use \emph{proof of work}. This defends the sybil attack.
But an incentive is needed for the nodes that expend their CPU resources.

@@

\begin{lemma}\label{lem:con_consistent}
  If a node sees a valid $\mathcal{C}_{r-1}$ and another node sees a valid
  $\mathcal{C}'_{r-1}$, then $\mathcal{C}_{r-1} = \mathcal{C}'_{r-1}$.
\end{lemma}
\begin{proof}
  (sketch) $\mathcal{C}_{r-1}$ is signed by at least $N-f$ promoters from round
  $r-1$.
\end{proof}

\begin{lemma}\label{lem:con_alive}
  Eventually all node sees a valid $\mathcal{C}_{r-1}.$  
\end{lemma}
\begin{proof}
  (sketch) Liveness is satisfied because $\mathcal{C}_{r-1}$ is eventually propagated to
  all node by gossiping.
\end{proof}

@@

\begin{lemma}\label{lem:promoter_consistent}
  The new set of promoter for the next consensus round is consistent with
  respect to all the nodes in the network.
\end{lemma}
\begin{proof}
  (sketch) All nodes use the same deterministic function to compute the luck
  value.
\end{proof}

Nodes should wait for some time to collect the CP blocks, so they wait for some
time $\Delta$ before moving on to the next phase. Note that promoters waiting
for a some time $\Delta$ to collect transactions does not violate the
asynchronous assumption because this behaviour can be seen as a long delay in
the asynchronous system.

\begin{corollary}
  Setup phase satisfies agreement and total order because the protocol is run
  deterministicly on the same input (\cref{lem:con_consistent},
  \cref{lem:promoter_consistent}). It also satisfies liveness due to
  \cref{lem:con_alive}.
\end{corollary}

@@

There are techniques that uses no dealers. First is to use
PBFT~\cite{castro1999practical}, but we must change our asynchronous assumption
into the weak synchrony assumption. Most likely this is not possible because of
the ``When to start?'' problem. It's difficult to give a bounded delay for
propagating the consensus result.

@@

\begin{lemma}
  Consensus phase satisfied agreement, total order and liveness.
\end{lemma}
\begin{proof}
  Defer proof? Refer to the papers.
\end{proof}

\begin{theorem}
  Checkpoint consensus satisfied agreement, total order and liveness.
\end{theorem}
\begin{proof}
  (sketch) Both phases are asynchronous so we do not need to make assumptions on
  when begin phase begins. Both phases also satisfy the agreement, total order
  and liveness.
\end{proof}

\section{Validation}\label{sec:validation}

Since we reach consensus on checkpoints rather than all the transactions, we
need to detect fraud, such as forks. Further, we need to ensure the system is
secure in a sense that valid transactions cannot be forged into invalid
transactions once it has reached consensus, and vice versa.

First, we define the requirements of a valid transaction $t_{i,j}$ as follows,
much of it is derived from the TrustChain model.

\begin{enumerate}
\item There exist a pair $t_{i', j'}$ that satisfies the pair definition.
\item $t_{i,j}$ and $t_{i', j'}$ is created using \texttt{newtx}, i.e. valid
  signatures and hash pointers, etc..
\item There exist an agreed piece that contains $t_{i, j}$ and another agreed
  piece that contains $t_{i', j'}$. All the blocks in the agreed pieces have
  valid hash pointers, the blocks in the pieces do not need to be valid.
\item The CP blocks $c_{i, k}, c_{i', k'}$ that immediately follow $t_{i, j}$
  and $t_{i', j'}$ are created using \texttt{newcp} using the same consensus
  result $\mathcal{C}_r$ as input and are in the agreed pieces.
\end{enumerate}

Blocks that do not satisfy the definition above are not necessarily invalid. It
may be the case that the validity cannot be determined due to incomplete
information. For such cases we say its validity is unknown.
Now we define an invalid transaction.

\begin{enumerate}
\item TODO
\end{enumerate}

Now we define the properties that are desired by the validation protocol.
\begin{enumerate}
\item \emph{Correctness}: The validation protocol outputs the correct result
  according to the aforementioned requirements.
\item \emph{Agreement}: If any correct node decides on the validity of a 
  transaction, then all other correct nodes are able to reach the same
  conclusion.
\item \emph{Liveness}: Any valid transactions can be validated eventually.
\item \emph{Unforgeability}: If some transaction is determined to be valid, then
  it cannot be changed to an invalid transaction, the opposite also applies.
\end{enumerate}

Note that the input of the validation protocol is a TX block, so these
properties hold with respect to TX blocks. In practice, if a node has a set of
TX blocks that are in the unknown state, then it must run the validation
protocol to determine whether they are valid. Further, these conditions do not
imply all invalid transactions (forks) can be found globally, only the validity
of the TX blocks that the honest nodes are interested in can be determined. The
advantage of this scheme is that it saves a lot of computational and bandwidth
costs because nodes only run the validation protocol on the TX blocks of their
own interest.

\subsection{Validation protocol}
Assume node $u$ is aware of all the past consensus results $\mathcal{C}_r$.
Suppose $u$ wish to validate $t_{i,j}$. It performs the following.

\begin{enumerate}
\item Determine the pair $t_{i', j'}$.
\item Find the agreed enclosure for $t_{i,j}$ and $t_{i', j'}$ from
  $\mathcal{C}_r$, otherwise return ``unknown''.
%\item Find the smallest $c_{i', b}$ such that $b > j'$ is in consensus.
\item Query $i$ and $i'$ for the agreed pieces and ensure hash pointers are
  correct. Otherwise return ``unknown''.
\item Check that $t_{i,j}$ and $t_{i', j'}$ are in the agreed pieces and are
  created correctly using \texttt{newtx}. Otherwise return ``invalid''.
\item Check the checkpoints $c_{i, k}$ and $c_{i', k'}$ that immediately follow
  $t_{i,j}$ and $t_{i', j'}$ are in the agreed pieces and are created in the
  same round, i.e. $\texttt{round}(c_{i, k}) = \texttt{round}(c_{i', k'})$.
  Otherwise return ``invalid''.
\item Return ``valid''.
\end{enumerate}

Most likely $u = i$ or $u = i'$, because they are incentivised to check the
validity of their TX blocks that are of unknown validity.

\subsection{Analysis}
In this section we analyse the validation protocol with respect to the four
properties in \cref{sec:validation}.

\begin{lemma}
  The validation protocol is correct.
\end{lemma}
\begin{proof}
  Correctness follows directly from the protocol specification, namely it
  directly implements the validation requirements.
\end{proof}

\begin{lemma}
  The validation protocol satisfies the agreement property.
\end{lemma}
\begin{proof}
  We proof by contradiction. Suppose two nodes $i$ and $j$ where $i \ne j$ runs
  the protocol on the same TX block $t_{k, l}$, $i$ outputs ``valid'' and $j$
  outputs ``invalid''. $i$ and $j$ would have both picked the same enclosure
  $c_{k, a}$ and $c_{k, b}$, as these are determined from the consensus result.
  Then $k$ must produce two agreed pieces that start with $c_{k,a}$ and end with
  $c_{k, b}$ with valid hash pointers, but one satisfies the validation protocol
  and the other one doesn't (e.g. missing $t_{k, l}$). Producing these two
  pieces is equivalent to producing a collision. Due to the properties of
  cryptographic hash functions this is not possible, thus a contradiction.
\end{proof}

\begin{lemma}
  The validation protocol does not satisfy the liveness property.
\end{lemma}
\begin{proof}
  (sketch) Nodes may be offline.
\end{proof}

There are many ways around the liveness. As long as the TX is validated once,
then the agreed pieces can be gossiped.

\begin{lemma}
  A valid TX block cannot be made into an invalid TX block and vice versa.
\end{lemma}
\begin{proof}
  We proof by contradiction. Suppose $t_{k, l}$ is determined to be valid using
  the enclosure $c_{k, a}$ and $c_{k, b}$ and the agreed piece $p$. Then $t_{k,
    l}$ is forged into an invalid transaction. To forge it into an invalid
  transaction the attacker must either tamper with the data within the block or
  remove $t_{k, l}$ from $p$. In other words create $p'$ where $p \ne p'$,
  enclosed by $c_{k, a}$ and $c_{k, b}$, and has valid hash pointers. Producing
  $p'$ is equivalent to finding a collision, thus a contradiction.

  TODO proof ``vice versa''.
\end{proof}

\subsection{Time and message complexity}
For a single transaction, the message complexity is linear with respect to the
size of the agreed pieces for both part of the TX pair. Time complexity is
constant since we only perfrom two queries (to $i$ and $i'$).

\section{Fraud detection}
Existence testing detects fraud when a single party of the TX is malicious. To
detect fraud when both parties are malicious, every node performs existence
testing on all TX blocks between CP blocks as if they are some other node. The
assignment is determined using CP block digest where every node is assigned to
the nearest node that has a higher digest, cycle back if there is no higher
digest.

If the number of malicious parties is low, then it's more probable that a fork
is detected. If where there are a large number of malicious parties, we repeat
the random sampling process.
